# DAY #1 (Workshops)

### Efficient R programming:

1. [Workshop materials](https://www.jumpingrivers.com/t/2018-erum-efficient/#/)
2. [Tutorial](https://github.com/jr-packages/efficientTutorial)

#### .Rprofile:

```{r remedy001}

options(prompt  = ":D",
        digits = 5,
        Continue = " ") # (hide "+")

```

```{r remedy002}

.First <- function() {
  cat("\nWelcome at", date(), "\n")
}
.Last <- function() {
  cat("Bye")
}

```
 
### Clean R code

1. Materials

- [Cheatsheet](https://github.com/czeildi/erum-2018-clean-r-code/blob/infra/clean_r_code_cheatsheet_2018-05-08.pdf)
- [code](https://github.com/czeildi/erum-2018-clean-r-code)

2. Suggestions:
- Function for every small functionality
- You do not always need to create one large  function (it depends)
- Function names should tell what tasks the accomplish, not what operations they perform, e.g inner_join -> wrapper add_country_metadata

3. Sourcing every script in a file:

```{r remedy003, eval = FALSE}

purrr::walk(list.files("R", full.names = TRUE))


```
4. [Materials](gitgitgithub.com/czeildi/erum_2018_clean_r_code)

5.

  shortcut |  action
  ------------- | -------------
  Alt  + - | assignment operator
  Ctrl + Alt + X | extract function from selection
  CMD + I | reindent lines




6. __Book: Robert C. Martin - Clean Code__

### Building package that lasts (Colin Fay)

1. [ThinkR](http://thinkr.fr)

2. Tools:

  - `desc #(editing description)`
  - `usethis` https://github.com/r-lib/usethis
  - `use_dev_package("callr")`
  - `use_rcpp()`
  - `use_test("my-test")`
  - `use_data(x, y)`
  - `use_tidy_description`
  - `use_build_ignore`
  - `available::available("ggplot2")`

3. Everything you did during engeneering: 
```{r remedy004}
usethis::use_data_raw -> devstuff.R

```

4. Version number:

- _major.minor.patch_
- start: 0.0.0.9000
- Increment +1 every bugfix

5. `use_build_ignore` - creates regexes to buildignore

6. Creating Dockerfiles for R code
  - [containerit project](https://github.com/o2r-project/containerit)
  - [containerit package](http://o2r.info/2017/05/30/containerit-package/)

7. Investigating conflicts:

```{r remedy005, echo = FALSE}

tidyverse::tidyverse_conflicts

```

8. Dev oriented optimisation

- "If copy something even twice - write a function"

9. User oriented optimisation

attempt package - stop_if_not, stop_if_any, stop_if_all

10. Benchmarking


```{r remedy006}

library(magrittr)
library(microbenchmark)
library(ggplot2)

microbenchmark(rchisq(100, 0),
               rchisq(100, 1),
               rchisq(100, 2),
               rchisq(100, 3),
               rchisq(100, 5), times=1000L) %>% autoplot()

```

New package: "bench"

11. Suggestion: use R formats - RDS instead of csv

12.

```{r remedy007, eval=FALSE}

skip_if_not(curl::has_internet())

```

13. Test with rhub:

```{r remedy008, eval = FALSE}

check_with_valgrind()

```

14 `Revdepcheck` - reverse dependency (for CRAN packages)

15. [Temporarilly modifying global state using withr](https://github.com/r-lib/withr)

16. [Addin for markdown reports](https://github.com/ThinkR-open/remedy)

18. [Shiny app for adding prefixes to functions](https://github.com/dreamRs/prefixer)


### Keras

https://rstudio.github.io/reticulate/

see scripts in rstudio (be attached later)

# DAY 2 (Conference)

### Rust code in R


- https://github.com/r-rust 
- https://github.com/r-rust/hellorust


Why to use?
- Safer, no pointers, memory leaks - it is memory safe
- As safe as C
- Native compiler for cross platform
- No need of install anything else except R pkg containing compiled Rust code
- Has version dependency
- Compiled binary does not depend on Rust or Cargo libs

Package [gifski](https://github.com/r-rust/gifski)


### OO systems in R (S3)

(generic.class/asDate.numer is equivalent to `summary.lm`

```{r remedy009, eval = FALSE}

# summary <- function(object, …)
useMethod("summary")

sum <- function(..., na.rm = FALSE)
	.Primitive("sum")

```



Dispatch:
`summary(lm(y~x))` → `summary.lm(lm(y~x))`ś

Why to use:
- Flexible & extensible
- Complex types can inherit after simpler type


Example:
```{r remedy010, results= FALSE}

dt <- data.table::data.table(mtcars)
print.data.frame(dt)
data.table:::print.data.table(dt)

```


TO check:

- www.github.com/wch/r-source
- Sloop pkg by Hadley

Slides:

- https://github.com/czeildi 
- https://github.com/czeildi/erum2018-oop-essentials


### Sentiment Analysis on Social Media and Big Data

- R pkg "iSAX"

### Taking inspiration from proven frontend frameworks to add to Shiny with 6 new packages (appsilon)

- [Shiny.semantic](https://github.com/Appsilon/shiny.semantic) let’s you download semantic UI CSS packages, fully customazible https://cran.r-project.org/web/packages/shiny.semantic/index.html 

- [Semantic.dashboard](https://github.com/Appsilon/semantic.dashboard) theme change easy (just put library(semantic.dashboard) instead of library(shiny.dashboard))
- [Shiny.router](https://github.com/Appsilon/shiny.router) makes routing possible with shiny (let’s you pass the args in url)
- [Shiny.router](https://github.com/Appsilon/shiny.router)  makes routing possible with shiny (let’s you pass the args in url)

Just for Appsilon clients:

- Shiny.users - access levels for users, authentication handled by shiny done on the app’s side. May be used with google account or ldap
- Shiny.admin - collect stats about app’s activity (what was clicked, how long the session took, etc.)

### Unified Content Analytics Process

- [Formattable package](https://github.com/renkun-ken/formattable) https://cran.r-project.org/web/packages/formattable/vignettes/formattable-data-frame.html 
- [rintrojs - interactive tutorials](https://github.com/carlganz/rintrojs)

www.ioki.pl/blog


### Erin LeDell: Scalable Automatic Machine Learning in R

https://github.com/h2oai/h2o-meetups/tree/master/2018_05_15_eRum_AutoML_in_R

### Szilard Pafka: Better than Deep Learning - Gradient Boosting Machines (GBM) in R

- [Package LightGMB](https://github.com/Microsoft/LightGBM/tree/master/R-package)
- [szilard/GBM-perf](https://github.com/szilard/GBM-perf)

### Andrie de Vries: Tools for using TensorFlow with R

[tfruns package](https://github.com/rstudio/tfruns)

`ls_run()` → look for hyperparameters:

Parameters can be flagged, e.g. number of runs can be set in list() and then best model can be chosen by ordering by modell accuracy

`Cloudml`

Books:
Deeep learning with R
Deeo Learning, Googfellow


### Robin Lovelace: Geocomputation for Active transport planning: a case study of cycle network design

https://cran.r-project.org/web/packages/robCompositions/index.html
https://github.com/cran/robCompositions


### Way to doing RNA-seq

https://github.com/federicomarini/erum2018

- R pkgs: "pcaExplorer", "ideal"
- iSEE - interactiveSummarizedExperimentExplorer


# DAY #3 (Conference)

### A practical history of R (where things came from)

- Henrik Bengtsson: A Future for R: Parallel and Distributed Processing in R for Everyone
- https://github.com/HenrikBengtsson/future.batchtools
- Shiny v1.1. Announced and will cointaint promises
- radii.defer: Deferred execution of nested functions

radii.cache

Will be available: https://github.com/tdeenes/radii.defer


### Barbara Borges: Drilldown data discovery with Shiny

https://github.com/bborgesr/erum2018


### Colin Gillespie: Getting the most out of GitHub and friends

R CMD check --as-cran mypkg.tar.gz

Travis
Appveyor: https://www.appveyor.com/


Arthur Charpentier: Demographics with Genealogical Data
https://3wen.github.io/genealogy/


### Simon Field: Exploiting Spark for high-performance scalable data engineering and data-science on Microsoft Azure

Graph computation on Spark with R: https://www-03.ibm.com/services/learning/ites.wss/zz-en?pageType=badges&id=4fbc7101-62cf-48de-b47d-c3b9e28e4824





### How to choose R package?

- Crantastic: https://www.crantastic.org/
- Metacran: https://www.r-pkg.org/
- Rweekly: https://rweekly.org/

Rmazing code
- http://style.tidyverse.org/
- [rOpenSci style guide](https://github.com/ropensci/onboarding/blob/master/packaging_guide.md)

### Robust data pipelines
[Drake R package](https://github.com/ropensci/drake)

Tomaszilagyi.com


### Achim Zeileis: R/exams -- A One-for-All Exams Generator

- https://arsnova.uibk.ac.at/mobile/#
- www.r-exams.org


### Tracking changes in data with the lumberjack package

https://github.com/data-cleaning/dcmodify 

Lumberjack R pkg to track changes in data

"daff" R package
Edwin de Jonge: validatetools - resolve and simplify contradictive or redundant data validation rules
https://github.com/data-cleaning/validatetools

### Why to use validate?

To create rule sets
Detect conflicting rules

```{r remedy011, eval = FALSE}

rules <- validator( if (age < 16) income == 0
                     , job %in% c("yes", "no")
                     , if (job == "yes") income > 0)

v <- confront(iris, rules)

quality <- as.data.frame(v)
merge(iris, quality)

```


### Andrea Melloncelli: What software engineers can teach to data scientists: code safety and … with automatic tests
"Shinytest" pkg

Testing upfront (Test Driven approach)

### Wit Jakuczun: Know your R usage workflow to handle reproducibility challenges

www.rsuite.io


### Omayma Said: Fitting Humans Stories in List Columns: Cases From an Online Recruitment Platform

```{r remedy012, eval = FALSE}

tidyr::nest()
tidyr::unnest()

```

- https://github.com/ropensci/solrium
- http://lucene.apache.org/solr/guide/7_3/solr-tutorial.html
- https://en.wikipedia.org/wiki/Apache_Solr
